---
title: |
  | **Tentative Topics Schedule**
  | **Applied Linear Algebra**
author: |
  | MTH 342 Section A
  | Spring 2024
  | University of Portland
bibliography: references.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(file.path("R","functions.R"))
```

```{r echo=FALSE}
create_footer()
```

<br>

> See Books & Online Resources Lists for the readings & practice materials.
>
> The reading materials are not mandatory but it is encouraged.

<br>The "Reading" column in the table below contains page numbers (Pg.) or chapters (ch.) on which it refers to a label in the Books & Online Resources List. For example "Pg. 1-5 [N]" refers to pages 1-5 of the first item in the list, which is the textbook titled "Linear algebra with applications".

<br>

## **Topics and Materials**

| **Week** | **Day** | **Topic** | **Worksheet** | **Homework** | **Reading** |
|:-----:|:-----:|:---------------:|:-----:|:-----:|:-----:|
| **1** | M 1/15 | _MLK Day_ | - | - | - |
|  | W 1/17 | _Cancelled_ | - | - | - |
|  | F 1/19 | Introduction and Orientation to ALA<br>_(Temporarily Asynchronous)_ | Algebra & Geometry Review | - | Syllabus |
| **2** | M 1/22 | Dot Product &<br>Projections | Vectors and the Geometry of Space Part 1 | - | Ch. 4.1 & 4.2 [N]  |
|  | W 1/24 | Cross Product &<br>Planes | Vectors and the Geometry of Space Part 2 | - | Ch. 4.3 [N] |
|  | F 1/26 | Equations of lines and Planes | Lines and Planes in Space | Assigned: Homework 1 | TBA |
| **3** | M 1/29 | Linear Independence &<br>Matrix Geometry | Basis Vectors and Independence | - | TBA |
|  | W 1/31 | Matrix Addition, Multiplication, and Transposition | Matrix Algebra and Applications | - | TBA |
|  | F 2/2 | System of Linear Equations & Row Reduction | System of Linear Equations and Applications | Assigned: Homework 2 | TBA |
| **4** | M 2/5 | Matrix Rank | - | - | TBA |
|  | W 2/7 | Matrix Inverse | Matrix Rank and Inverse | - | TBA |
|  | F 2/9 | Elementary Matrices | Elementary Matrices | - | TBA |
| **5** | M 2/12 | Linear Transformations | TBA | - | TBA |
|  | W 2/14 | _Project Period_ | - | - | - |
|  | F 2/16 | **Project Phase 1** | - | Assigned: Homework 3 | - |
| **6** | M 2/19 | Vector Spaces | TBA | - | TBA |
|  | W 2/21 | Subspace and Spanning Sets | TBA | - | TBA |
|  | F 2/23 | Orthogonality | TBA | Assigned: Homework 4 | TBA |
| **7** | M 2/26 | Eigentheory | TBA | - | TBA |
|  | W 2/28 | Real and Complex Eigenvalues | TBA | - | TBA |
|  | F 3/1 | Eigenvectors | TBA | Assigned: Homework 5 | TBA |
| **8** | M 3/4 | _Spring Vacation_ | - | - | - |
|  | W 3/6 | _Spring Vacation_ | - | - | - |
|  | F 3/8 | _Spring Vacation_ | - | - | - |
| **9** | M 3/11 | Eigenvectors for Distinct Eigenvalues | TBA | - | TBA |
|  | W 3/13 | Eigenvectors for Repeated Eigenvalues | TBA | - | TBA |
|  | F 3/15 | Diagonalization | TBA | - | TBA |
| **10** | M 3/18 | Matrix Factorizations | TBA | - | TBA |
|  | W 3/20 | _Project Period_ | - | - | - |
|  | F 3/22 | **Project Phase 2** | - | Assigned: Homework 6 | - |
| **11** | M 3/25 | LU Factorization | TBA | - | TBA |
|  | W 3/27 | QR Factorization | TBA | - | TBA |
|  | F 3/29 | _Easter Vacation_ | - | Assigned: Homework 7 | - |
| **12** | M 4/1 | _Easter Vacation_ | - | - | - |
|  | W 4/3 | Singular Vectors | TBA | - | TBA |
|  | F 4/5 | Singular Value Decomposition (SVD) | TBA | Assigned: Homework 8 | TBA |
| **13** | M 4/8 | Linear and Non-Linear Least Squares | TBA | - | TBA |
|  | W 4/10 | Least Squares Classification | TBA | - | TBA |
|  | F 4/12 | Principal Components | TBA | - | TBA |
| **14** | M 4/15 | Principal Component Analysis (PCA) | TBA | - | TBA |
|  | W 4/17 | _Project Period_ | - | - | - |
|  | F 4/19 | **Project Phase 3** | - | Assigned: Homework 9 | - |
| **15** | M 4/22 | _Project Period_ | TBA | - | TBA |
|  | W 4/24 | _Project Period_ | TBA | - | TBA |
|  | F 4/26 | _Project Period_ | TBA | Assigned: Homework 10 (Optional) | TBA |
| **16** | W 5/1 | **Project Phase 4** | - | - | - |

<br>

Along with textbook [N] and [B], most of the course materials (contents of worksheets and homework) of each topic was taken from these following sources:

  * Inquiry oriented linear algebra (IOLA) by @iola2013
  * Introduction to linear algebra (6th Edition) by @strang2023linearalgebra
  * Linear algebra and learning from data by @strang2019linearalgebradata
  * Linear algebra by @linearalgebramitopencourseware

<br>

## **Books & Online Resources Lists**

```{r echo=FALSE}
refs <- bibtex::read.bib("references.bib")
```

*Click on the link to access the resources.*

### Textbooks

[N] `r capture.output(print(refs["nicholson2018linear"]))` [[pdf]](https://lyryx.com/wp-content/uploads/2018/01/Nicholson-OpenLAWA-2018A.pdf){target="_blank"}

[B] `r capture.output(print(refs["boyd2018introduction"]))` [[pdf]](https://web.stanford.edu/~boyd/vmls/vmls.pdf){target="_blank"}

<br>

\newpage

## **References**

::: {#refs}
:::

<br>
